---
title: "Limpet Egg Analysis"
author: "Pascal Bärtschi"
output:
  html_document:
    theme: simplex
    toc: true
    toc_float: true
    toc_depth: 5
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

In this analysis, we'll explore the relationships between limpet egg count and factors such as density and season. We'll begin by loading the necessary libraries and importing the dataset. Let's dive in!

```{r}
# Importing libraries
suppressPackageStartupMessages({
  # Importing libraries
  library("ggplot2")
  library("dplyr")
  library("ggfortify")
  library('tidyverse')
  library('GGally')
})


# Load data
# setwd("~/BSc_UZH/UZH_22FS/BIO144/datasets-master")
data <- read.csv('data/limpet.csv', stringsAsFactors = T)
```

### Data Exploration

We'll start by getting a glimpse of the dataset to understand its structure and contents.

```{r}

glimpse(data)
```

Next, we'll generate summary statistics to gain insights into the key features of the dataset.


```{r}

summary(data)
```

### Visualizing Relations

Let's visualize the relationship between egg count, density, and season using a scatter plot.

```{r}

ggplot(data, aes(x = DENSITY, y = EGGS, colour = SEASON)) + 
  geom_point() + 
  scale_color_manual(values = c(spring = "green", summer = "red")) +
  theme_bw()
```

### Model Construction

We'll construct a linear regression model to understand how egg count is influenced by density and season.

```{r}

model_limp <- lm(EGGS ~ DENSITY * SEASON, data = data)
```

### Model Check and Interpretation

Let's take a closer look at the model and its components.

```{r}
autoplot(model_limp, smooth.colour = NA)
```

### ANOVA Table

An analysis of variance (ANOVA) will help us understand the significance of different factors in our model.

```{r}

anova(model_limp)
```

### Model Summary

The summary of the linear regression model provides important insights into the relationships we're exploring.

```{r}

summary(model_limp)
```

### Coefficient Table

The coefficients of the model reveal specific impact levels of the factors on egg count.

```{r}

coef(model_limp)
```

### Predictions and Visualization

We'll use the model to make predictions and visualize the results.

```{r}

new_x <- expand.grid(DENSITY = seq(from = 8, to = 45, length.out = 10),
                     SEASON = levels(data$SEASON))

predEGG <- predict(model_limp, newdata = new_x, interval = 'confidence')

addThese <- data.frame(new_x, predEGG) 
addThese <- rename(addThese, EGGS = fit)

EGGatMeanDens <- data.frame(new_x, predEGG)
EGGatMeanDens <- rename(EGGatMeanDens, EGGS = fit)

ggplot(data, aes(x= DENSITY, y = EGGS, colour = SEASON)) + 
  geom_point(size = 5) +
  geom_smooth(data = addThese, 
              aes(ymin = lwr, ymax = upr, fill = SEASON), stat = 'identity') +
  scale_color_manual(values = c(spring = 'green', summer = 'red')) +
  scale_fill_manual(values = c(spring = 'green', summer = 'red')) +
  theme_bw()

```

### Conclusion and Interpretation

In conclusion, our analysis provides insights into the relationships between egg count, density, and season in limpets. The visualizations and statistical analysis allow us to better understand these dynamics.