---
title: "Using multiple linear regression to analyse bodyfat"
author: "Pascal Bärtschi"
output:
  html_document:
    toc: true
    toc_depth: 5
output_dir: "../html"
date: "`r Sys.Date()`"
---
# Earthworm Analysis

```{r setup, include=FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(ggfortify)
library(tidyverse)
library(GGally)
```

### Data Import and Preprocessing

```{r}

# Load data
data <- read.csv('data/earthworm.csv', stringsAsFactors = TRUE)
```

### Exploratory Data Analysis

```{r}

# Summary statistics
summary(data)

# Unique values
unique(data)

# Frequency table of 'Gattung'
table(data$Gattung)
```

We start by importing the data and performing initial exploratory data analysis.

```{r}

# Histogram of 'Gewicht'
ggplot(data, aes(x = Gewicht)) +
  geom_histogram()
```

The histogram of 'Gewicht' variable shows this distribution.

```{r}

# Log transformation of 'Gewicht'
data <- mutate(data, log10_Gewicht = log10(Gewicht))

# Histogram of log-transformed 'Gewicht'
ggplot(data, aes(x = log10_Gewicht)) +
  geom_histogram(bins = 20)
```

We apply a log transformation to 'Gewicht' for less skewed distribution.

```{r}

# Scatterplot of 'Magenumf' vs. 'Gewicht'
ggplot(data, aes(x = Magenumf, y = Gewicht)) +
  geom_point()
```

The scatterplot of 'Magenumf' against 'Gewicht' shows a non-linear relationship.

```{r}

# Scatterplot of 'Magenumf' vs. log-transformed 'Gewicht'
ggplot(data, aes(x = Magenumf, y = log10_Gewicht)) +
  geom_point()
```

The scatterplot with the log-transformed 'Gewicht' shows a stronger linear relationship.

### Linear Regression

```{r}

# Linear regression model
m1 <- lm(log10_Gewicht ~ Magenumf, data)
```

We build a linear regression model (m1) to predict log-transformed 'Gewicht' using 'Magenumf'.

```{r}

# Visualization of the model
autoplot(m1)
summary(m1)
```

We visualize the model and provide a summary. We see that the residuals seem randomly distributed, as well as the data is not skewed (QQ)

```{r}

# Scatterplot with linear regression line
ggplot(data, aes(x = Magenumf, y = log10_Gewicht)) +
  geom_point() + 
  geom_smooth(method = 'lm') +
  xlab('Gut Circumference [mm]') +
  ylab('Worm Weight [g]')
```

The scatterplot with the linear regression line visualizes the relationship between 'Magenumf' and log-transformed 'Gewicht'.

### Comparison Among Genera

```{r}

# Scatterplot of 'Gattung' vs. log-transformed 'Gewicht'
ggplot(data, aes(x = Gattung, y = log10_Gewicht)) +
  geom_point() + 
  geom_boxplot()
```

We compare log-transformed 'Gewicht' among different genera using scatterplots and boxplots.

```{r}

# Linear model with 'Gattung' as explanatory variable
m2 <- lm(log10_Gewicht ~ Gattung, data)
autoplot(m2)
summary(m2)
anova(m2)
```

We build a linear model (m2) using 'Gattung' as an explanatory variable and provide its summary and ANOVA results. We see that `Gattung` seems to be important in explanatory power, thus the type of worm is important for prediction.

### Multiple Explanatory Variables

```{r}

# Scatterplot with color-coded 'Gattung'
ggplot(data, aes(x = Magenumf, y = log10_Gewicht, colour = Gattung)) +
  geom_point()
```

We visualize the relationship between 'Magenumf' and log-transformed 'Gewicht', color-coded by 'Gattung'.

```{r}

# Multiple variable linear regression model
m3 <- lm(log10_Gewicht ~ Magenumf + Gattung, data)
autoplot(m3)
summary(m3)
```

We build a linear regression model (m3) using both 'Magenumf' and 'Gattung' as explanatory variables and provide its summary. This model enables 3 intercepts for each worm but only one slope.

```{r}

# Linear regression model with interaction effects
m4 <- lm(log10_Gewicht ~ Magenumf * Gattung, data)
autoplot(m4)
summary(m4)
```

We build a model (m4) that includes interaction effects between 'Magenumf' and 'Gattung' and provide its summary. This enables independent slopes of the categories as visualized below: 

### Visualization of Model Results

```{r}

# Scatterplot with color-coded 'Gattung' and linear regression lines
ggplot(data, aes(x = Magenumf, y = log10_Gewicht, colour = Gattung)) +
  geom_point() + 
  geom_smooth(method = 'lm') +
  xlab('Gut Circumference [mm]') +
  ylab('Worm Weight [g]')
```

The scatterplot with color-coded 'Gattung' and linear regression lines with indpendent slopes visually summarizes the model results.