---
title: "Using multiple linear regression to analyse bodyfat"
author: "Pascal Bärtschi"
output:
  html_document:
    theme: simplex
    toc: true
    toc_float: true
    toc_depth: 5
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
# Load necessary libraries
library(tidyverse)
library(ggfortify)
library(GGally)
```

### Data Import and Preprocessing

```{r}

# Load data
dd <- read_delim('data/bodyfat.txt', '\t', escape_double = FALSE, trim_ws = TRUE, show_col_types = FALSE)

# Select only three columns
dd <- select(dd, bodyfat, weight, abdomen)
```

### Distribution Overview

```{r}
# Plot histograms of each variable
dd %>% 
  gather(key = "measurement", value = "value") %>% 
  ggplot() + 
  geom_histogram(aes(x = value)) + 
  facet_wrap(~measurement, scales = "free")
```

The histograms show the distribution of the bodyfat, weight, and abdomen variables.

### Exploratory Data Analysis

```{r}

# Visualize pairwise relationships
ggpairs(dd)

# Linear Regression Model
m1 <- lm(bodyfat ~ weight + abdomen, data = dd)
```

The pairwise scatterplots provide insights into the relationships between the variables. We then build a linear regression model (m1) to predict bodyfat using weight and abdomen as explanatory variables.

```{r}

# Visualize residuals
autoplot(m1)

# Summary of the model
summary(m1)
```

The scatterplot of residuals helps us assess the model's assumptions. The summary provides information about the coefficients and statistical significance of the model.

```{r}

# Confidence intervals for coefficients
confint(m1)
```

The confidence intervals for the model coefficients provide insights into the precision of the estimates.

### Model Validation

```{r}

# Predicted values and residuals
predicted_values <- predict(m1)
fitted_values <- fitted(m1)

# Visualize predicted vs. actual values
ggplot(data.frame(predicted = predicted_values, actual = dd$bodyfat), 
       aes(x = actual, y = predicted)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Actual Bodyfat", y = "Predicted Bodyfat")
```

We validate the model by plotting predicted versus actual values. The scatterplot and regression line help us assess the model's predictive performance.

### Model Improvement

```{r}

# Generate new data for visualization
new_data <- expand.grid(weight = mean(dd$weight),
                        abdomen = seq(min(dd$abdomen), max(dd$abdomen), length = 100))

# Predictions with new data
predictions <- predict(m1, newdata = new_data, interval = 'confidence')
predictions_df <- cbind(new_data, predictions)

# Visualize predicted bodyfat vs. abdomen
ggplot(predictions_df) + 
  geom_line(aes(x = abdomen, y = fit)) +
  geom_smooth(mapping = aes(x = abdomen, y = fit, ymin = lwr, ymax = upr),
              stat = 'identity') +
  geom_point(data = dd, aes(x = abdomen, y = bodyfat))
```

We improve the model by generating new data for visualization and predicting bodyfat values based on abdomen and weight measurements, which both seems to be accurate predictors of bodyfat

### Model Comparison and Interpretation

```{r}
# Reload to get more data
dd <- read_delim('data/bodyfat.txt', '\t', escape_double = FALSE, trim_ws = TRUE, show_col_types = FALSE)
# Select only  columns
dd <- select(dd, bodyfat, weight, abdomen, ankle, wrist, height)

dd %>% 
  gather(key = "measurement", value = "value") %>% 
  ggplot() + 
  geom_histogram(aes(x = value)) + 
  facet_wrap(~measurement, scales = "free")
```


```{r}
# Create models for comparison
m_both <- lm(bodyfat ~ weight + abdomen, data = dd)
m_weight <- lm(bodyfat ~ weight, data = dd)
m_abdomen <- lm(bodyfat ~ abdomen, data = dd)

# Calculate adjusted R-squared values
r2both <- summary(m_both)$adj.r.squared
r2weight <- summary(m_weight)$adj.r.squared
r2abdomen <- summary(m_abdomen)$adj.r.squared

# Calculate unique variance for each variable
r2_ua <- r2both - r2weight
r2_uw <- r2both - r2abdomen

# Calculate shared explanatory power
r2_shared <- r2both - r2_ua - r2_uw

r2 <- c(r2_ua, r2_uw, r2_shared)
names(r2) <- c("abdomen only", "weight only", "shared")
print(r2)
```

We compare three models (m_both, m_weight, and m_abdomen) to understand the individual and shared explanatory power of weight and abdomen on bodyfat. We see that weight exlaines the least variance in the data.

### Variable Scaling and Interpretation

```{r}

# Mutate variables for scaling
dd <- mutate(dd, scaled_abdomen = scale(abdomen), scaled_weight = scale(weight))

# Build model with scaled variables
m_scaled <- lm(bodyfat ~ scaled_abdomen + scaled_weight, data = dd)

# Compare scaled model with original model
summary(m_scaled)
summary(m_both)
```

We standardize variables to prevent misinterpretations due to different scales and build a model (m_scaled) using scaled variables.

### Broom Package for Model Summaries

```{r}

# Use broom package for model summaries
library(broom)
tidy(m_scaled)
augment(m_scaled)
glance(m_scaled)
```

The broom package provides convenient functions (tidy, augment, glance) to extract useful information from the model.

### Visualizing Model Coefficients

```{r}

# Mutate additional variables for scaling
dd <- mutate(dd, scaled_wrist = scale(wrist), scaled_ankle = scale(ankle), 
             scaled_height = scale(height))

# Build model with additional scaled variables
m <- lm(bodyfat ~ scaled_height + scaled_ankle + scaled_wrist + scaled_abdomen + scaled_weight, data = dd)

# Extract coefficient information
td <- tidy(m, conf.int = TRUE)

# Visualize coefficients
ggplot(td, aes(estimate, term)) +
  geom_point() +
  geom_errorbar(mapping = aes(xmin = conf.low, xmax = conf.high), width = 0.1 )
```

We visualize the coefficients of the model using scatterplots with error bars.

### Conclusion

In this analysis, we performed exploratory data analysis, built linear regression models, improved model fit, and compared different models to understand the relationship between bodyfat, weight, and abdomen. We explored various aspects of regression analysis and model interpretation to gain insights into the variables' influence on body fat percentages.