---
title: "Model Selection with dredge: An Echocardiogram Analysis"
author: "Your Name"
output:
  html_document:
    toc: true
    toc_depth: 5
output_dir: "../html"
date: "`r Sys.Date()`"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Importing Libraries

We load the necessary libraries that will be used throughout the analysis.

```{r}

suppressPackageStartupMessages({
  library("ggplot2")
  library('tidyverse')
  library("MuMIn")
})
```

### Load Data

We import the echocardiogram dataset for our analysis.

```{r}

data <- read.csv('data/echocardiogram.csv', na.strings = c('?'))
```

### Data Cleaning and Overview

We perform data cleaning steps to handle missing values and rename columns for clarity.

```{r}

data <- na.omit(data)

names(data) <- c("survival", "still_alive", "age_at_heart_attack", "pericardial_effusion",
                  "fractional_shortening", "epss",
                  "lvdd", "wall_motion_score", 
                  "wall_motion_index", "mult",
                  "name",
                  "group",
                  "alive_at_1")

# data <- data %>% select(-name)

glimpse(data)

```

### Model Fitting

We fit a logistic regression model to predict survival status based on various echocardiogram measurements.

```{r}

glm1 <- glm(alive_at_1 ~ fractional_shortening +
              epss +
              lvdd +
              wall_motion_index +
              pericardial_effusion, data = data, family = binomial, na.action = 'na.fail')
```

### Model Selection

We perform model selection using the dredging technique and select the best-fitting model based on likelihood ratios.

```{r}

model_dredge <- dredge(glm1)
best_dredge <- get.models(model_dredge, subset = delta == 0)[[1]]
# Perform Likelihood Ratio Test
anova(best_dredge, test = 'LRT')
```

### Model Summary

Here, we provide a summary of the selected model's coefficients and statistical information.

```{r}

summary(best_dredge)
```

### Visualizing Predictions

We visualize the predicted probabilities of survival based on the wall motion index.

```{r}

ggplot(data, aes(x = wall_motion_index, y = alive_at_1)) +
  geom_point(position = position_jitter(height = 0.1, width = 0), alpha = 0.5) +
  stat_smooth(method = "glm", method.args = list(family = "binomial")) +
  theme_bw()
```

### ANOVA for Full Model

We conduct an analysis of variance (ANOVA) for the full model to assess the overall significance of predictor variables.

```{r}

anova(glm1, test = 'Chisq')
```
