---
title: "Ruggedness and GDP Analysis"
author: "Pascal Bärtschi"
output:
  html_document:
    toc: true
    toc_depth: 5
params:
  output_dir: "../html"
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../../../')

```

### Introduction

In this analysis, we will investigate the relationship between ruggedness, GDP, and the continent of Africa. We'll explore data transformations, visualizations, and use ANCOVA to determine whether there is a significantly different relationship between GDP and ruggedness in countries within and outside of Africa. Let's get started!
Clear Brain and Setup

We'll start by loading necessary libraries and importing the data set.

```{r}
# Importing libraries
suppressPackageStartupMessages({
  # Importing libraries
  library("ggplot2")
  library("dplyr")
  library("ggfortify")
  library('tidyverse')
  library('GGally')
})


# Load data

data <- read.csv('00-data/rugged.csv', stringsAsFactors = T)
```

### Data Exploration

Let's begin by visualizing the distribution of the rgd_ppc_2000 variable using a histogram.

```{r}

ggplot(data, aes(x= rgdppc_2000)) +
  geom_histogram()
```

To deal with the skewed distribution, we'll apply a logarithmic transformation to the variable.

```{r}

data <- mutate(data, rgd_log10 = log10(rgdppc_2000))

ggplot(data, aes(x= rgd_log10)) +
  geom_histogram()
```

### Creating New Variables

We'll create a new categorical variable cont_africa1 based on the conf_africa variable to distinguish between African and non-African countries.

```{r}

data <- mutate(data, cont_africa1 = as.factor(ifelse(data$cont_africa == 1, 'Africa', 'non Africa')))
# Check distribution of countries
table(data$cont_africa1)
```

### Ruggedness and Continent Relationship

Let's explore the distribution of ruggedness and its potential relationship with the continent.

```{r}

ggplot(data, aes(x= rugged)) +
  geom_histogram()
```

Next, we'll identify the country with the lowest ruggedness.

```{r}

data %>%
  arrange(desc(rugged)) %>%
  select(country)
```

### Handling Missing Values

We'll remove missing values and create a new dataset for further analysis.

```{r}

sum(is.na(data$rgd_log10))
datar <- na.omit(select(data, rugged, cont_africa1, rgd_log10))

# Check distribution of continent in the new dataset
table(datar$cont_africa1)
```

### Visualizing Relationships

Let's visualize the relationship between ruggedness, log-transformed GDP, and the continent using a scatter plot.

```{r}

ggplot(datar, aes(x = rugged, y = rgd_log10, color = cont_africa1)) +
  geom_point() + 
  geom_smooth(method = 'lm')
```

### ANCOVA Analysis

We'll conduct an ANCOVA analysis to examine the significance of the relationship between GDP and ruggedness, accounting for the effect of the continent.

```{r}

m1 <- lm(rgd_log10 ~ rugged * cont_africa1, data = datar)
summary(m1)
autoplot(m1, smooth.colour = NA)
```

### Distribution of Residuals

Let's visualize the distribution of residuals to assess the assumptions of the ANCOVA model.

```{r}

resid <- residuals(m1)

ggplot(m1, aes(x = residuals(m1))) +
  geom_histogram(bins = 12)
```

### Predictions and Visualization

We'll generate predictions using the ANCOVA model and visualize the results.

```{r}

new.x <- expand.grid(rugged = seq(min(datar$rugged), max(datar$rugged), length.out = 170),
                     cont_africa1 = levels(datar$cont_africa1))

pred_gdp <- predict(m1, newdata = new.x, interval = 'confidence')

addThese_gdp <- data.frame(new.x, pred_gdp)
addThese_gdp <- rename(addThese_gdp, rgd_log10 = fit)

ggplot(datar, aes(x = rugged, y = rgd_log10, colour = cont_africa1)) +
  geom_point() +
  geom_smooth(data = addThese_gdp, aes(ymin = lwr, ymax = upr, fill = cont_africa1),
              stat = 'identity')
```

### ANOVA Analysis

Finally, let's perform ANOVA to draw conclusions about the significance of the relationships we've explored.

```{r}

anova(m1)
```

### Conclusions

In conclusion, our ANCOVA analysis reveals that there is a significantly different relationship between GDP and ruggedness in countries within and outside of Africa (p-value = 0.0032). This implies that the continent has a significant influence on this relationship. These findings provide valuable insights into the interactions between economic and geographical factors.