---
title: "Analyzing Algal Growth and Herbivore Interactions in Intertidal Zones: two way ANOVA"
author: "Pascal Bärtschi"
output: 
  html_document:
    toc: true
    toc_depth: 5
output_dir: "../html"
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../../../')
```


This document explores the interaction between herbivores and height in relation to algal cover. We will analyze a dataset to uncover potential patterns and relationships using various data analysis techniques.

### Prerequisites and Data Loading

Importing the necessary libraries and loading the data set.

```{r}
suppressPackageStartupMessages({
  # Importing libraries
  library("ggplot2")
  library("dplyr")
  library("ggfortify")
  library('tidyverse')
  library('GGally')
})


# data
data <- read.csv("00-data/intertidalalgae.csv", stringsAsFactors = T)  
```

### Data Exploration

We'll begin by getting an overview of the dataset using the glimpse() and summary() functions. Additionally, we'll examine the levels of the "herbivores" variable.

```{r}
glimpse(data)
```

```{r}
summary(data)
```

```{r}
table(data$herbivores)
```


```{r}
levels(data$herbivores)
```

### Group Summary

We will group the data by "height" and "herbivores" and calculate summary statistics.

```{r}
summary_data <- data %>%
  group_by(height, herbivores) %>%
  summarise()
summary_data
```

### Distribution of Algal Area

Let's visualize the distribution of algal area using a histogram, grouped by "height" and "herbivores".

```{r}
ggplot(data, aes(x = Area_cm2)) +
  geom_histogram() +
  facet_grid(~height ~ herbivores) +
  theme_bw()
```

### Interaction Visualization

First, we'll transform the area variable using the square root and update the dataset.

```{r}
data <- mutate(data, Area_root = sqrt(Area_cm2))
```

And second, we'll create scatter plots to visualize the interaction between "herbivores" and "height" with respect to the transformed algal area.

```{r}
ggplot(data, aes(x = height, y = Area_root, color = herbivores)) +
  geom_point(position = position_jitterdodge(jitter.width = 0.2))
```

```{r}
ggplot(data, aes(x = height, y = Area_root, color = herbivores)) +
  geom_boxplot()
```

### Summary Statistics of Algal Cover

Compute the summary statistics of algal cover and standard errors.
```{r}
summary_algae <- data %>%
  group_by(height, herbivores) %>%
  summarise(mean = mean(Area_root), sd = sd(Area_root), se = sd(Area_root) / sqrt(n()))

options(pillar.sigfig = 4)
summary_algae

```

### Interaction Modeling

We will use linear regression to model the interaction between "herbivores" and "height" on the algal cover.

```{r}
model_algae <- lm(Area_root ~ herbivores * height, data = data)
```

The `autoplot()` function serves to check skewedness of the data or correlations between fitted values and residuals.

```{r}
autoplot(model_algae, smooth.colour = NA)
```

### ANOVA and Model Summary

We perform ANOVA and summarize the model.

```{r}
anova(model_algae)
```

```{r}
summary(model_algae)
```

### Confidence Intervals

We calculate confidence intervals for the model coefficients.

```{r}
conf_intervals <- confint(model_algae, level = 0.95)
conf_intervals
```

### Interaction Plot with Confidence Intervals

We create a plot of the interaction along with confidence intervals for interpretation.

```{r}

ggplot(summary_algae, aes(x = height, y = mean,
                   color = herbivores, group = herbivores)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.1) +
  ylab(expression("Algal cover - square root cm"^"2")) +
  xlab('Height') +
  theme_bw()
```

### Conclusion

In this analysis, we explored the interaction between herbivores and height with regards to algal cover. The interaction plot and model results provide insights into the relationship between these variables.


