---
title: "Mixed Effects Model Analysis"
author: "Your Name"
output:
  html_document:
    toc: true
    toc_depth: 5
output_dir: "../html"
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../../../')
```

### Introduction

In this document, we will analyze the behavior of mixed effects models using the lme4 library in R. Each section will focus on specific tasks related to mixed effects models, such as model fitting, hypothesis testing, and exploring influential points. Interpretation of function outputs and results will be provided through comments in the code.

```{r}

# Load the lme4 library
suppressPackageStartupMessages({library(lme4)})
```

### Data Loading

Let's start by loading the dataset using the read.csv() function. The dataset contains information about politeness data.

```{r}

# Read in the dataset
politeness <- read.csv("http://www.bodowinter.com/tutorial/politeness_data.csv")
```

### Exploratory Data Analysis

Before diving into model analysis, let's perform some exploratory data analysis to understand the structure of the dataset.

```{r}

# Display column names
print(paste0("Colnames:", colnames(politeness)))

# Display summary of data
print(str(politeness))
print(summary(politeness))
```

### Missing Data Handling

We'll address missing values in the dataset using various methods.

```{r}

# Check for missing values and their locations
which(is.na(politeness$frequency))
```

```{r}
# Alternative method to check for missing values
which(!complete.cases(politeness))
```

### Relationship Visualization

To explore relationships, we'll create a boxplot to visualize the relationship between frequency, attitude, and gender.

```{r}

# Visualize the relationship using a boxplot
boxplot(frequency ~ attitude * gender, col = c('white', 'lightgray'), data = politeness)
```

### Fitting Mixed Effects Models

We'll fit mixed effects models to analyze the relationship between variables.

#### Model 1: Random Effects

Fitting a mixed effects model with random effects for subjects and scenarios.

```{r}

# Fit a mixed effects model with random effects
lmm1 <- lmer(frequency ~ attitude + (1|subject) + (1|scenario),
             data = politeness)
summary(lmm1)
```

#### Model 2: Including Gender

Fitting a mixed effects model including the gender variable.

```{r}

# Fit a mixed effects model with gender and random effects
lmm2 <- lmer(frequency ~ attitude + gender + (1|subject) + (1|scenario),
             data = politeness)
summary(lmm2)
```

### Hypothesis Testing

We'll perform hypothesis testing using the anova() function to compare models and assess significance.

```{r}

# Perform hypothesis testing using anova
null_lmm <- lmer(frequency ~ gender + (1|subject) + (1|scenario),
                 data = politeness, REML = F)

full_lmm <- lmer(frequency ~ attitude + gender + (1|subject) + (1|scenario),
                 data = politeness, REML = F)

anova(null_lmm, full_lmm)
```

### Interaction Effects

We'll investigate interaction effects and their significance using different model configurations.

```{r}

# Testing interaction effects using anova
full_inter <- lmer(frequency ~ attitude * gender + (1|subject) + (1|scenario),
                 data = politeness, REML = F)

full_plus <- lmer(frequency ~ attitude + gender + (1|subject) + (1|scenario),
                 data = politeness, REML = F)

full_1 <- lmer(frequency ~ 1 + (1|subject) + (1|scenario),
                 data = politeness, REML = F)

anova(full_inter, full_plus, full_1)
```

### Coefficient Interpretation

Let's explore the coefficients of the model to understand their implications.

```{r}

# Explore the coefficients of the model
coef(full_lmm)
```

### Different Slopes Model

We'll analyze a model with different slopes to account for individual variability.

```{r}

# Fit a model with different slopes
lmm3 <- lmer(frequency ~ attitude + gender + (1+attitude|subject) + (1+attitude|scenario),
             data = politeness, REML = F)
coef(lmm3)
```

### Influential Points

We'll identify influential points using a loop and visualization.

```{r}

# Identify influential points using a loop
all.res = numeric(nrow(politeness))
for (i in 1:nrow(politeness)) {
  myfullmodel = lmer(frequency ~ attitude + (1+attitude|subject), politeness[-i,])
  all.res[i] = fixef(myfullmodel)[2]
}
```

### Conclusion

In this analysis, we delved into mixed effects model analysis using the `lme4` library in R. We covered data loading, exploratory data analysis, missing data handling, model fitting, hypothesis testing, interaction effects, coefficient interpretation, and the identification of influential points. Through these sections, we demonstrated the process of fitting and interpreting mixed effects models for real-world data analysis.

