---
title: "Analyzing Daphnia Growth and Parasitism: Uncovering Relationships with one way ANOVA"
author: "Pascal Bärtschi"
output: 
  html_document:
    toc: true
    toc_depth: 5
output_dir: "../html"
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



This code explores the relationship between Daphnia growth and parasitism using data analysis techniques. Daphnia, tiny aquatic organisms, are indicators of ecosystem health. We delve into a dataset called "Daphniagrowth.csv" to investigate potential associations between growth rate and the presence of parasites.

#### Prerequisites and data loading
```{r include = FALSE}
suppressPackageStartupMessages({
  # Importing libraries
  library("ggplot2")
  library("dplyr")
  library("ggfortify")
  library('tidyverse')
  library('GGally')
})

# load data
data <- read.csv("data/Daphniagrowth.csv", stringsAsFactors = TRUE)

```


#### Data Exploration

We begin by loading the data set and gaining initial insights using `glimpse()` and `summary()`. We also create a distribution of parasite occurrences with the `table()` function.

```{r}
glimpse(data)
```


```{r}
summary(data)
```


```{r}
table(data$parasite)
```

#### Visual Insight

Visualizations play a vital role. We use ggplot2 to create a box and whisker plot, showing growth rate distribution across parasite groups. Flipping the axes clarifies the view.

```{r}
ggplot(data, aes(x = parasite, y = growth.rate)) + 
  geom_violin(fill = "orange", color = "orange",  alpha = 0.05) +
  geom_boxplot(fill = NA, color = "darkgreen") +
  theme_bw() + 
  coord_flip() # flips axes to see full parasite names 
```


#### Modeling the Connection

Linear regression is employed to model growth rates based on parasitism. The model, named model_grow, estimates growth variations among parasite groups.

```{r}
model_grow = lm(growth.rate ~ parasite, data = data)
```



#### Model Visualization

The `autoplot()` function serves to check skewedness of the data or correlations between fitted values and residuals.

```{r}
autoplot(model_grow, smooth.color = NA)
```
The data is a bit skewed (Q-Q) but seem ok, however we need to note that the standardized residuals show tiny correlation (Scale-Location) with the fit.

#### Statistical Analysis

We perform ANOVA using `anova()` to assess if parasite groups lead to distinct growth rates. The F-value informs us about between-group versus within-group variance.

```{r}
anova(model_grow)
# H0: all groups come from populations with the same mean
# F-value: ratio of in between group variance to the within group variance
```
* The "parasite" factor has 3 degrees of freedom and a sum of squares of 3.1379. The mean sum of squares is 1.04597,        resulting in a relatively high F value of 32.325.
* The associated p-value (Pr(>F)) is very close to zero (2.571e-10), which is much smaller than the typical significance    level of 0.05. This suggests that the effect of the "parasite" factor on the "growth.rate" is statistically significant.


```{r}
summary(model_grow)
```

The intercept is representative for the control group. Thus, we estimate of growth of p.r: 0.6, p.p: 1.05, m.b: 0.85
on estimated 36 degrees of freedom with this linear model.

#### Group Comparisons

Using `summarise()`, we compute mean growth rates for each parasite group. The results are visualized alongside individual data points, offering a comprehensive growth rate trend overview.



```{r}
sumDat <- data %>%
  group_by(parasite) %>%
  summarise(meanGR = mean(growth.rate))

sumDat
```



#### ANOVA visualization: grouped means of growth by parasite

```{r}
ggplot() +
  geom_point(data = data, mapping = aes(x= parasite, y = growth.rate, color = parasite)) +
  geom_point(data = sumDat, mapping = aes(x = parasite, y = meanGR, color = parasite), size= 3) + 
  theme_bw() +
  coord_flip()

```
#### Conclusion

This ANOVA analysis unravels that there is a relationship between parasite and growth rate of daphnia, indicated by the relatively high F-value and the p-value close to zero.




